---
title: "Session2(9-1)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Helpful Tips: 
option+command+i (shortcut to paste R section)

```{r - Helpful Tips}
```
### Vectors
Prints out a charactor vector
```{r Character vectors}
vec1 <- c("a", "b", "c", "d", "e")
print(vec1)
```
Prints out a numeric vector
```{r Numeric vector}
vec2 <- c(3.14159, 1.414, 5.678, 9.01) #comment
print(vec2)

```

#### Matrix
Matrix = 2-D vector 
```{r Matrix}
m1 <- matrix(c(3.14159, 1.414, 5.678, 9.01),
               nrow=2,
               ncol=2)
print(m1)
             
```

```{r Matrix v2}
m2 <- matrix(data = vec2, ##using vec2 variable 
               nrow=2,
               ncol=2)
print(m2)
```
```{r Identical method}
identical(m1,m2)
```

### Data frame
Collection of individual vectors put together in a 2x2 vector. Vectors HAVE TO BE THE SAME LENGTH in order to do so (i.e. # of elements inside of a vector)
```{r Intro to Data frame}
d1 <- data.frame(name = c("Jake", "Abby", "Bailey", "Xavier"), 
                 age = c(11,20,19,5), 
                 male = c(TRUE, FALSE, FALSE, TRUE))
print(d1)
```
#### Subset a data frame 

Print first column of d1 
```{r}
d1 # also prints d1 without print method
print(d1$name) #print first column

```
```{r}
print(d1[1,1]) #row first, column second
print(d1[1, ])#comma is important
```
Print column names of the data frame
```{r}
names(d1) # variables names in the data frame 
```
Print the first two rows and the first two columns of d1
```{r}
print(d1[c(1,2), c(1,2)]) #vectors being combined. since data frames are composed of vectors 
```
```{r}
print(d1[1:2, 1:2]) # another version 
```
Print 1st and3rd row and the 1st and 3rd columns
```{r}
print(d1[c(1,3), c(1,3)])
```
### List
```{r Named list}
list1 <- list(v1 = vec1, v2 = vec2, mat1 = m1, df = d1) #dont print lists because it has a dataframe, vector everything combined together and it is a MESS! 
print(list1) #prints entire list
print(list1[c(1,3)]) #prints only the 1st and 3rd of list.
```


